Synthesizing Slow Motion Periodic Videos via Circular Coordinate Reshuffling And Voting

### 2018-01-17 Paper/paper.tex
Wrote a little over half of the paper text now

% Template for ICIP-2018 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}

% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}

% Title.
% ------
\title{AUTHOR GUIDELINES FOR ICIP 2018 PROCEEDINGS MANUSCRIPTS}
%
% Single address.
% ---------------
\name{Author(s) Name(s)\thanks{Thanks to XYZ agency for funding.}}
\address{Author Affiliation(s)}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
Abstract
\end{abstract}
%
\begin{keywords}
One, two, three, four, five
\end{keywords}
%

>>>>

%%TODO: IF THIS GETS IN, CITE MY THESIS (since I present an early version
%of this idea in my background and copied some text from there)

% Template for ICIP-2018 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,amssymb,amsfonts,amsthm,graphicx}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx

\graphicspath{{Figures/}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\title{Synthesizing Slow Motion Periodic Videos via Circular Coordinate Reshuffling And Voting}

%
% Single address.
% ---------------
\name{Author(s) Name(s)\thanks{Thanks to XYZ agency for funding.}}
\address{Author Affiliation(s)}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
\begin{document}
%\ninept
%
\maketitle
%

\begin{abstract}
Abstract
\end{abstract}
%
\begin{keywords}
One, two, three, four, five
\end{keywords}
%

\section{Introduction}

\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{1DExample.pdf}
\caption{An illustration of our technique on a 1D periodic time series with additive Gaussian noise ($n_t \in \mathcal{N}(0, 1)$).  We estimate the period to be 24 (a multiple of the fundamental frequency), and perform a 24-length sliding window embedding.  From this, we use the mean of the birth and death time of the largest persistence dot (row 2, column 2) to autotune the spatial bandwidth for the graph Laplacian (upper right two figures for the weighted and unweighted Laplacian), from which we derive a map to the circle (middle, right two plots).  From this, we derive circular coordinates, which we can use to do a simple reordering (row 3 column 2) or median consensus (row 3, column 3), using interpolated votes from all windows interleaved by the coordinates (row 3, column 2). }
\label{fig:ConceptFigure}
\end{figure}

In this work, we aim to find slow motion representatives for videos which have periodic motion.  The result should loop seamlessly and to occur at a higher framerate than the captured framerate, allowing a smooth, high detail view of a single cycle of the motion.  In the absence of a high framerate camera, recent work has shown it is possible to boost the framerate of an ordinary camera by inserting a coded aperture in front of the lens \cite{llull2013coded}.  By contrast, in our work, we take a purely signal processing approach which avoids any specialized hardware.

Our approach is {\em Eulerian}; that is, we process the video pixel by pixel with no tracking.  Unlike Fourier-based Eulerian motion video processing tools (\cite{wu2012eulerian, wadhwa2013phase}), however, we take a {\em geometric} approach by quantifying {\em sliding window embeddings} of our videos (Section~\ref{sec:slidingwindow}).  Sliding window embeddings of periodic time series form samples of a topological loop \cite{perea2015sliding}, as do generalized multivariate sliding windows on periodic video data \cite{traliehigh, tralie2017quasi}.  This implies that, at the right scale, a nearest neighbor graph built on the sliding window point cloud should be approximately a circle graph, even if the overall extrinsic geometry is quite complicated.  Bearing this in mind, we use Laplacian eigenmaps \cite{belkin2003laplacian} of this embedding to construct maps from these point clouds to the circle, or {\em circular coordinates} (Section~\ref{sec:laplacian}).  In particular, each sliding window is assigned a phase $\phi \in [0, 2\pi]$.  To autotune the scale at which Laplacian eigenmaps are built, we use topological data analysis \cite{edelsbrunner2000topological,edelsbrunner2008persistent,edelsbrunner2010computational,carlsson2009topology,ghrist2014elementary} to automatically find the scale at which the most prominent loop in the data occurs.

Once we have extracted a phase for each window, we use this information to reorder the content of the video.  We also exploit the sliding window structure to have different windows vote on the final pixels in each frame of the template (Section~\ref{sec:cyclereordering}).  As an illustration, Figure~\ref{fig:ConceptFigure} shows an example of our technique on a sampled periodic time series (i.e. single pixel grayscale video).  There are only 12 samples per period in the original signal, so the details of each period coarse and noisy.  However, once they are re-sorted, we get a nice, fine-detailed representation of one period.  The extension of this technique to videos will be described in detail in Section~\ref{sec:methods}.

\section{Releated Work}
Our technique unifies many disparate ideas that have appeared before in different contexts, in addition to the references we mentioned in the introduction.  Sliding window embeddings, or ``delay reconstructions,'' have found a diverse array of applications in activity recognition \cite{frank2010activity,venkataraman2016shape}, gene expression data \cite{perea2015sw1pers}, EEG analysis \cite{stam2005nonlinear, plesnik2014detection}, audio and music analysis \cite{herzel1994analysis,serra2009cross,bello2011measuring,traliemoebius}, video analysis \cite{schodl2000video, tralie2017quasi}, and motion capture analysis \cite{venkataraman2016shape}.

The graph Laplacian has been used to re-arrange images around a loop as a pre-processing step for structure from motion \cite{averbuch2015ringit}, and vector diffusion maps have been used to re-order the frames of microscope images a developing embryo \cite{dsilva2015diffusionvecordering}.  A different topological approach with cohomology circular coordinates\cite{de2011persistent} was used to parameterize a sliding window embedding of the Lorenz attractor \cite{de2012topological} and periodic activites in motion capture data \cite{vejdemo2015cohomological}.  Neural networks have also been trained to find circular coordinates \cite{levy2015live,anafi2017cyclops}, though this is no longer unsupervised.

There has also been some related work on sliding window embeddings to synthesize ``video textures,'' or videos that repeat dynamics indefinitely \cite{schodl2000video}, and there has been work synthesizing seamless video loops by clustering pixels into different periods and adjusting them to loop through a common period \cite{Liao2013VideoLoops,Liao2015VideoLoops}.

\section{Methods}
\label{sec:methods}

\begin{algorithm}[t]
  \caption{Sliding Window Video Loops}\label{alg:videoreordering}
  \begin{algorithmic}[1]
    \Procedure{ReorderVideo}{$X$, weighted, median} \\
    \Comment{$X$ is a $W \times H$ video $X$ with $N$ frames}
    \State $\hat{X} \gets$ 1D ISOMAP(X)
    \State $d \gets$ fundamental period estimate of $\hat{X}$
    \State $Y \gets SW_{d}[X[n]] \in \mathbb{R}^{WHd}$, with $M = N-d+1$ frames
    \State Perform SVD $Y = USV^T$, $\hat{Y} \gets V^TY$
    \State $I \gets dgm_1[\hat{Y}] = ((b_1, d_1), ... (b_k, d_k))$
    \State $i \gets \argmax_{i}[d_i - b_i]$, $\sigma \gets b_i + \frac{1}{2}(d_i - b_i)$
    \State $d_{ij} \gets ||\hat{Y}_i - \hat{Y}_j||_2$
    \If {weighted}
        \State $A_{ij} \gets e^{-d_{ij}^2/\sigma}$
    \Else
        \State $A_{ij} \gets \left\{ \begin{array}{cc} 1 & d_{ij} \leq \sigma \\ 0 & \text{otherwise} \end{array} \right\}$
    \EndIf
    \State $A_{ii} \gets 0, D_{ii} \gets \sum_{j=1}^M A_{ij}, D_{i, j \neq i} \gets 0$, $L \gets D - A$
    \State $v_1, v_2 \gets $ adjacent eigenvectors of $L$ with fewest zero-crossings
    \State $\theta \gets \cos^{-1}(v_2/v_1)$
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

We now describe our algorithm in several stages.  Throughout our discussion, we assume we are dealing with 3 channel color videos.  For a frame resolution of $W \times H$, each frame can be thought of as a Euclidean vector in $\mathbb{R}^{W \times H \times 3}$.  Algorithm~\ref{alg:videoreordering} shows the full algorithm.

\subsection{Sliding Window Embeddings}
\label{sec:slidingwindow}

Given a $W \times H$ video $X(t) \in \mathbb{R}^{3WH}$ indexed by time, the sliding window video embedding \cite{cao1998dynamics,traliehigh,tralie2017quasi} is defined as

\begin{equation}
SW_{d, \tau}[X(t)] = \left[ \begin{array}{c} X(t) \\ X(t + \tau) \\ \vdots \\ X(t + (d-1)\tau)  \in \mathbb{R}^{3WHd} \end{array} \right]
\end{equation}

where $\tau$ is a step size and $d$ is the dimension of the embedding.  In this work, we fix $\tau = 1$ for simplicity and drop $\tau$ in the subscript, and we also sample times $t$ only at integer frame indices.  In theory, a sliding window embedding of a single generic observation function of a dynamical system of intrinsic dimension $m$ can be reconstructed \cite{takens1981detecting}.
Talk about fundamental frequency estimation \cite{Mcleod05asmarter}

\subsection{Laplacian Eigenmaps And Persistent Homology}
\label{sec:laplacian}

which means the first two nonzero eigenvectors should be a sine and cosine, or orthogonal linear combinations therein.  When plotted against each other, they make an approximate circle, and the arctangent of the two eigenvector coordinates at every window can be used to determine the {\em phase} of the corresponding window in the periodic signal.  The first sample of each window can then be re-sorted by phase.

Talk about autotuning with peristent homology, mention that it's the opposite of the approach in \cite{bendich2011improving}.

\subsection{Cycle Reordering}
\label{sec:cyclereordering}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\columnwidth]{XTSlice.pdf}
\caption{An XT slice of a line of pixels (magenta line, upper left) over time for the original video and for reordered videos with and without median consensus.}
\label{fig:XTSlice}
\end{figure}

Figure~\ref{fig:XTSlice} shows the difference between a simple reordering and a median consensus reordering.  Due to natural variation from cycle to cycle, the simple reordering has many temporal discontinuities when interleaving these cycles.  By contrast, the median voting is clean, and it has the added benefit of removing nonperiodic background components.

\subsection{Sharpening Postprocessing}

The results of the median recycling are smooth and qualitatively similar to the results in the ``motion denoising'' work by the authors of \cite{rubinstein2011motion} on time lapse videos, though both results tend to be slightly blurry.

\section{Experiments}

Robustness experiments, jumping jacks 2 men, two exercise videos from \cite{levy2015live}, beating heart videos from \cite{traliehigh}, \cite{wu2012eulerian}, and \cite{wadhwa2013phase}.

\section{Discussion}

Since we rely on the theory and constructions in \cite{tralie2017quasi}, we are subject to similar constraints in the allowable motion and drift for our method to work well.  We note that Eulerian video magnification also degrades in the presence of too much drift \cite{wu2012eulerian, wadhwa2013phase}.
